---
title: "Assignment_2"
author: "Yeswanth Siripurapu"
date: "2023-10-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Assignment Problem Statement

  Universal bank is a young bank growing rapidly in terms of overall customer
  acquisition. The majority of these customers are liability customers
  (depositors) with varying sizes of relationship with the bank. The customer
  base of asset customers (borrowers) is quite small, and the bank is 
  interested in expanding this base rapidly in more loan business. In particular
  , it wants to explore ways of converting its liability customers to personal
  loan customers.  
  
  A campaign that the bank ran last year for liability customers showed a
  healthy conversion rate of over 9% success. This has encouraged the retail
  marketing department to devise smarter campaigns with better target marketing.
  The goal is to use k-NN to predict whether a new customer will accept a loan
  offer. This will serve as the basis for the design of a new campaign.
  
  The file UniversalBank.csv contains data on 5000 customers. The data include
  customer demographic information (age, income, etc.), the customer’s 
  relationship with the bank (mortgage, securities account, etc.), and the 
  customer response to the last personal loan campaign (Personal Loan). Among 
  these 5000 customers, only 480 (= 9.6%) accepted the personal loan that was 
  offered to them in the earlier campaign.
  
  Partition the data into training (60%) and validation (40%) sets.
  
## Load required libraries

```{r}
library(class)
library(caret)
library(e1071)
library(knitr)
```

### Read the UniversalBank data

```{r}
universal.df <- read.csv("D:/UniversalBank.csv")
dim(universal.df)
t(t(names(universal.df))) # The t function creates a transpose of the data frame
```
### Drop ID and ZIP
```{r}
universal.df <- universal.df[,-c(1,5)]
```

### Transform categorical variables into dummy variables

```{r}
# Only Education needs to be converted to factor

universal.df$Education <- as.factor(universal.df$Education)

```

```{r}
# Now, convert Education to Dummy Variables

groups <- dummyVars(~., data = universal.df) # This creates the dummy groups

universal_m.df <- as.data.frame(predict(groups,universal.df))
```

### Split the data to 60% training and 40 % Validation 

```{r}
set.seed(1)  # Important to ensure that we get the same sample if we rerun the
#code
train.index <- sample(row.names(universal_m.df), 0.6*dim(universal_m.df)[1])
valid.index <- setdiff(row.names(universal_m.df), train.index)  
train.df <- universal_m.df[train.index,]
valid.df <- universal_m.df[valid.index,]
t(t(names(train.df)))

```

### Now normalize the data

```{r}
train.norm.df <- train.df[,-10] # Note that Personal Income is the 10th variable
valid.norm.df <- valid.df[,-10]

norm.values <- preProcess(train.df[, -10], method=c("center", "scale"))
train.norm.df <- predict(norm.values, train.df[, -10])
valid.norm.df <- predict(norm.values, valid.df[, -10])
```

### Question

    1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, 
    Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities
    Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember 
    to transform categorical predictors with more than two categories into dummy
    variables first. Specify the success class as 1 (loan acceptance), and use
    the default cutoff value of 0.5. How would this customer be classified?

# We have converted all categorical variables to dummy variables

## Let's create a new sample
```{r}

new_customer <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1)
```

### Normalize the new customer

```{r}
new.cust.norm <- new_customer
new.cust.norm <- predict(norm.values, new.cust.norm)

```

### Now let us predict using K-NN(k- Nearest neighbors)

```{r}

knn.pred1 <- class::knn(train = train.norm.df, 
                       test = new.cust.norm, 
                       cl = train.df$Personal.Loan, k = 1)
knn.pred1

```

    2. What is a choice of k that balances between overfitting and ignoring the
    predictor information?
### Calculate the accuracy for each value of k    
### Set the range of k values to consider

```{r}

accuracy.df <- data.frame(k = seq(1, 20, 1), overallaccuracy = rep(0, 20)) 
for(i in 1:20) 
  {knn.pred <- class::knn(train = train.norm.df, 
                         test = valid.norm.df, 
                         cl = train.df$Personal.Loan, k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.pred,as.factor(valid.df$Personal.Loan)
                                       ,positive = "1")$overall[1]
}
which(accuracy.df[,2] == max(accuracy.df[,2])) 
plot(accuracy.df$k,accuracy.df$overallaccuracy, main = "Accuracy Vs K",
     xlab = "k", ylab = "accuracy")

```

  3.Show the confusion matrix for the validation data that results from using 
  the best k.

### Confusion Matrix using best K=3

```{r}

knn.pred <- class::knn(train = train.norm.df,
                         test = valid.norm.df, 
                         cl = train.df$Personal.Loan, k = 3)

confusionMatrix(knn.pred,as.factor(valid.df$Personal.Loan))

```

    4.Consider the following customer: Age = 40, Experience = 10, Income = 84,
    Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, 
    Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
    Card = 1. Classify the customer using the best k.

# Load new customer profile

```{r}
new_customer2<-data.frame(
  Age = 40, 
  Experience = 10, 
  Income = 84, 
  family =2, 
  CCAvg = 2, 
  Education_1 = 0,
  Education_2 = 1, 
  Education_3 = 0, 
  Mortgage = 0, 
  Securities.Account = 0, 
  CDAccount = 0, 
  Online = 1, 
  CreditCard = 1)

```


```{r}

knn.pred1 <- class::knn(train = train.norm.df, 
                       test = new.cust.norm, 
                       cl = train.df$Personal.Loan, k = 3)
knn.pred1
```

## Print the predicted class (1 for loan acceptance, 0 for loan rejection)

```{r}

print("This customer is classified as: Loan Rejected")

```

    5. Repartition the data, this time into training, validation, and test sets
    (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Compare 
    the confusion matrix of the test set with that of the training and 
    validation sets. Comment on the differences and their reason.
    

### Split the data to 50% training and 30% Validation and 20% Testing 

```{r}
set.seed(1)
Train_Index1 <- sample(row.names(universal_m.df), 0.5*dim(universal_m.df)[1])
Val_Index1 <- sample(setdiff(row.names(universal_m.df),Train_Index1),
                     0.3*dim(universal_m.df)[1])
Test_Index1 <-setdiff(row.names(universal_m.df),union(Train_Index1,Val_Index1))
Train_Data <- universal_m.df[Train_Index1,]
Validation_Data <- universal_m.df[Val_Index1,]
Test_Data <- universal_m.df[Test_Index1,]
```

### Now normalize the data

```{r}
train.norm.df1 <- Train_Data[,-10]
valid.norm.df1 <- Validation_Data[,-10]
Test.norm.df1  <-Test_Data[,-10]

norm.values1 <- preProcess(Train_Data[, -10], method=c("center", "scale"))
train.norm.df1 <- predict(norm.values1, Train_Data[,-10])
valid.norm.df1 <- predict(norm.values1, Validation_Data[,-10])
Test.norm.df1 <-predict(norm.values1,Test_Data[,-10])

```

### Now let us predict using K-NN(k- Nearest neighbors)

```{r}

validation_knn = class::knn(train = train.norm.df1, 
                           test = valid.norm.df1,  
                           cl = Train_Data$Personal.Loan, 
                           k = 3)

test_knn = class::knn(train = train.norm.df1, 
                     test = Test.norm.df1,  
                     cl = Train_Data$Personal.Loan, 
                     k = 3)

Train_knn = class::knn(train = train.norm.df1, 
                     test = train.norm.df1,  
                     cl = Train_Data$Personal.Loan, 
                     k = 3)
```

### Validation confusion Matrix

```{r}
validation_confusion_matrix = confusionMatrix(validation_knn, 
                                      as.factor(Validation_Data$Personal.Loan), 
                                               positive = "1")

validation_confusion_matrix

```
### Test confusion Matrix

```{r}

test_confusion_matrix = confusionMatrix(test_knn, 
                                         as.factor(Test_Data$Personal.Loan), 
                                         positive = "1")


test_confusion_matrix

```

```{r}

Training_confusion_matrix = confusionMatrix(Train_knn, 
                                            as.factor(Train_Data$Personal.Loan), 
                                               positive = "1")

Training_confusion_matrix
```


# Difference

##Test vs.Train:

***Accuracy:*** Train has a higher accuracy (0.9772) compared to Test (0.9507).

**Reason:** This because of differences in the datasets used for evaluation.
Train may have a more balanced or easier-to-predict dataset.

***Sensitivity (True Positive Rate):*** Train has higher sensitivity (0.7589)
compared to Test (0.5875).

**Reason:** This indicates that Train's model is better at correctly identifying
positive cases (e.g., loan acceptances). It may have a lower false negative rate.

***Specificity (True Negative Rate):*** Train has higher specificity (0.9987)
compared to Test (0.99403).

**Reason:** This suggests that Train's model is better at correctly identifying
negative cases (e.g., loan rejections). It may have a lower false positive rate.

***Positive Predictive Value (Precision):*** Train has a higher positive 
predictive value (0.9827) compared to Test (0.92157).

**Reason:** Train's model is more precise in predicting positive cases, 
resulting in fewer false positive predictions.

## Train vs.Validation:

**Accuracy:** Train still has a higher accuracy (0.9772) compared to Validation
(0.958).

***Reason:*** Similar to the comparison with Test, Train may have a more 
balanced or easier-to-predict dataset.

**Sensitivity (True Positive Rate):** Train has higher sensitivity (0.7589) 
compared to Validation (0.625).

***Reason:*** Train's model is better at correctly identifying positive cases.
This indicates that Validation's model may have a higher false negative rate.

**Specificity (True Negative Rate):** Train has higher specificity (0.9987)
compared to Validation (0.9934).

***Reason:*** Train's model is better at correctly identifying negative cases. 
Validation's model may have a slightly higher false positive rate.

**Positive Predictive Value (Precision):** Train still has a higher positive 
predictive value (0.9827) compared to Validation (0.9091).

***Reason:*** Train's model is more precise in predicting positive cases, 
resulting in fewer false positive predictions.

## Potential Reasons for Differences:

**Data set Differences** Variations in the composition and distribution of data 
between different sets can significantly impact model performance. For  
illustration, one data set may be more imbalanced, making it harder to  
prognosticate rare events.   

**Model Variability** Differences in model configurations or  arbitrary 
initialization of model parameters can lead to variations in performance.

**Hyperparameter Tuning** Different hyper parameter settings,  similar as the 
choice of k in k- NN or other model-specific parameters, can affect model 
performance.

**Data unyoking** If the data sets are  resolve else into training,  
confirmation, and test sets in each evaluation, this can lead to variations 
in results, especially for small data sets.

**Sample Variability** In small data sets, variations in the specific samples 
included in the  confirmation and test sets can  impact performance  criteria .

**Randomness** Some models,  similar as neural networks, involve randomness in
their optimization process, leading to slight variations. 
